# Homework 2 - Hilde Younce

```{r}
# install.packages("rstan", repos = c('https://stan-dev.r-universe.dev', getOption("repos")))
library(ggplot2)
library(rstan)
```

## Problem 1:

```{r}
# Get data 
wine_train <- read.csv('whitewine-training-ds6040.csv')
```

```{r}
# Density plot 1: residual.sugar
ggplot(wine_train, aes(x=residual.sugar)) + geom_density(fill = "lightblue")
```

The density plot for residual.sugar does not look very normally distributed. There is a higher density around -1 that declines and then zeros out at around 2.5, and is skewed towards the right. 

```{r}
# Density plot 2: total.sulfur.dioxide
ggplot(wine_train, aes(x=total.sulfur.dioxide)) + geom_density(fill = "lightblue")
```

The density for total.sulfur.dioxide looks much more normal than the density plot for residual.sugar. There looks to be a mean value around -0.5, with a slight skew to the right. 

```{r}
# 1) 
variances <- apply(wine_train, 2, var)
print(variances)
```

```{r}
# fixed.acidity:
rstan_options(auto_write = TRUE)

N = length(wine_train$fixed.acidity)
y = wine_train$fixed.acidity
sigma = var(wine_train$fixed.acidity)
mu0 = 0
tau = 1
data_list = list(N=N, y=y, sigma=sigma, mu0=mu0, tau=tau)

results = stan(file='homework2.stan', data=data_list)
extract(results)
```

