# HW 4 - Hilde Younce

**Honor Pledge:** On my honor as a student, I have neither given nor received unauthorized aid on this assignment.

```{r message = FALSE, warning = FALSE}
library(brms)
library(ggplot2)
library(dplyr)
sales_data <- read.csv('sales-ds6040.csv')
```

```{r}
head(sales_data)
```

### Problem Statement

We aim to investigate how regional store sales are influenced by the neuroticism and conscientiousness scores of store managers. Specifically, we seek to understand differences in sales between food and coffee orders. To achieve this, we will implement a Bayesian hierarchical model that accounts for clustering by store ID and incorporates a linear regression framework to explore these sales differences. Our goal is to uncover insights into the role of manager personality in driving sales performance and identify stores expected to perform well even in the absence of highly neurotic or conscientious managers.

### Approach:

We will employ a Bayesian hierarchical model, focusing on a random slope mixed-effects design. This approach is ideal for our analysis as it accounts for clustering by store while allowing variability in sales trends across stores. Specifically, we will model sales as the outcome, with predictors for food versus coffee transactions, standardized neuroticism and conscientiousness scores, and interactions between these predictors. By incorporating store-level random effects, our model effectively captures differences in sales behavior across locations.

Using this hierarchical structure, we achieve the flexibility of performing store-specific regressions while leveraging the full dataset. This approach enables us to identify patterns and trends in sales across the region, offering a cohesive view of how manager personality traits impact store performance.

The model is as follows:

$$
Sales_{ij} = \beta_{0j} + \beta_{1j} * Food + \beta_{2j} * Con + \beta_{3j}*Neur + \beta_{4j} * Con * Food + \beta_{5j}*Neur*Food + \epsilon_{i}
$$

Which is equivalent to:

$$
y_{ij} = \beta_{0j} + \beta_{1j}x_{1ij} + \beta_{2j}*x_{2ij} + \beta_{3j}*x_{3ij} + \beta_{4j}*x_{4ij} + \beta_{4j}*x_{5ij} + \epsilon_i
$$

$$
\epsilon_i \sim N(0, \sigma^2) 
$$ $$
\beta_{0j} \sim N(\mu_{\beta_0}, \sigma_0^2)
$$ $$
\beta_{1j} \sim N(\mu_{\beta_1}, \sigma_1^2)
$$ Where:

$Sales_{ij}$ - Sales for month $i$ in store $j$

$\beta_{0j}$ - Random intercept for store $j$

$\beta_{1j}$ - Effect of food transactions for store $j$

$\beta_{2j}$ - Effect of conscientiousness on coffee sales for store $j$

$\beta_{3j}$ - Effect of neuroticism on coffee sales for store $j$

$\beta_{4j}$ - Difference in $\beta_{2j}$ effect on food sales for store $j$

$\beta_{5j}$ - Difference in $\beta_{3j}$ effect on food sales for store $j$

$\epsilon_{i}$ - Residual error

### Prior Selection and Rationale:

Our priors for this model are:

$$
\mu_{\beta_0}, \mu_{\beta_1} \sim N(0,10)
$$

$$
\sigma^2, \sigma_0^2, \sigma_1^2 \sim Half-Cauchy(0,2) 
$$

For our mu values, we are using a Normal distribution. We will choose uninformative hyperparameters since we do not have a strong prior belief about how the data should behave. We let 0 be the mean because we have no prior belief if the direction of the effect will be positive or negative. We let 10 be our variance to allow for a wide range of values.

For the sigma values, we are using a Half-Cauchy distribution. Again we will choose uninformative hyperparamters. We set 0 as the mean since we are using Half-Cauchy, and thus only looking at positive values. Our scale parameter is 2, which favors smaller variances and makes the distribution centered closer to zero.

### Findings:

```{r message = FALSE, warning = FALSE}
# brms model 
formula <- bf(sales ~ con + neur + food + con*food + neur*food + (food|store))
priors <- c(
  set_prior("normal(0, 10)", class = "b"),
  set_prior("cauchy(0,2)", class = "sd")
)
model <- brm(formula, data=sales_data, prior=priors, refresh=0, silent=2)
```

```{r}
summary(model)
```

**How does conscientiousness and neuroticism impact the sales of coffee and food, and are coffee and\
food impacted differently?**

To understand how store manager personality traits influence food and coffee sales, we interpret the regression coefficients from our model.

Coffee sales:

-   The coefficient for conscientiousness is 0.34, indicating that coffee sales increase by 0.34 transactions for every one-unit increase in conscientiousness.

-   Conversely, the coefficient for neuroticism is -0.41, meaning coffee sales decrease by 0.41 transactions for every one-unit increase in neuroticism.

-   These results suggest that conscientiousness has a small positive impact on coffee sales, while neuroticism has a slightly negative effect.

Food sales:

The interaction terms con \* food and neur \* food represent the offset from the coffee coefficients. To interpret their impact on food transactions, we sum the coffee coefficients and their corresponding interaction terms:

-   For conscientiousness: 1.01+0.34=1.35. This indicates that food sales increase by 1.35 transactions for every one-unit increase in conscientiousness.

-   For neuroticism: 0.11−0.41=−0.30. This shows that food sales decrease by 0.30 transactions for every one-unit increase in neuroticism.

Overall, conscientiousness has a significant positive impact on food sales, while neuroticism has a small negative effect.

Comparing food and coffee sales:

To investigate how food and coffee sales differ, we focus on the food and interaction coefficients:

-   The food coefficient (−0.84) reflects the baseline difference in sales between coffee and food, indicating that coffee tends to sell 0.84 more units than food on average.

-   The con \* food coefficient (1.01) suggests that conscientiousness has a much stronger positive effect on food sales compared to coffee sales.

-   The neur \* food coefficient (0.11) indicates that neuroticism has a slightly greater impact on food sales than coffee sales, but the difference is relatively small.

In summary, conscientiousness appears to have a more substantial positive influence on food sales compared to coffee, while neuroticism generally has a negative impact on both types of sales but with less differentiation.

**Once you control for the personality characteristics of the store managers, what stores should be\
performing well? (i.e. the rest of the employees might be great, but the store manager might be bringing sales down)**

```{r}
ranef_data <- ranef(model)$store
intercepts <- ranef_data[, , "Intercept"]

intercepts_df <- as.data.frame(intercepts)
intercepts_df <- intercepts_df %>%
  mutate(store = rownames(intercepts_df),
         lower = Estimate - 1.96 * Est.Error,
         upper = Estimate + 1.96 * Est.Error)

ggplot(intercepts_df, aes(x = Estimate, y = store)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.2) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Random Intercepts by Store ID",
    x = "Random Intercept Estimate",
    y = "Store ID")
```

In this plot, we are measuring the random intercept estimates and confidence intervals for each store. These random intercepts represent deviations in store performance from the average sales, after controlling for the fixed effects (including manager personality characteristics). Thus, stores with the highest random intercept estimates are expected to perform well when accounting for store manager personality, and low random intercept estimates indicate a store is not performing well in comparison to average sales. Based on our results, we see that stores 14, 12, 13, and 17 are performing best. On the other hand, stores 16, 9, and 11 are performing worst in comparison to average sales.

### Summary

After running a random slope Bayesian hierarchical model, we were able to draw the following insights:

-   In general, an increase in conscientiousness has a positive impact on both coffee and food sales.

-   Conscientiousness has a greater positive impact on food sales than coffee sales by a significant amount.

-   In general, an increase in neuroticism has a negative impact on food and coffee sales, although the magnitude of this impact is less than that of conscientiousness.

-   Neuroticism does not impact food and coffee sales differently by a significant amount.

-   After controlling for store manager personality, we expect stores 12, 14, and 17 to perform the best.

-   We expect stores 9, 11, and 16 to perform the worst. These stores might warrant further investigation into management or operational issues.

### Diagnostics

**Trace Plots:**

```{r}
mcmc_plot(model, type = "trace")
```

**Density Plots:**

```{r}
mcmc_plot(model, type = "dens_overlay")
```

**Coefficient Plots:**

```{r}
mcmc_plot(model, type = "intervals", pars = "b_")
```

![]()
